Turnover of species and interactions

### 2018-10-29 paper/data-exploration/turnover.Rmd
higher dpi in figures
Remove tables, improve figures and add text explanation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

@@

We start by checking wether species (and interactions) occur in enough localities to be able to detect a change across the species ranges.

@@

### Species turnover

@@

plot_freq_step <- function(freq_df){
  if ("guild" %in% names(freq_df)){
    p <- freq_df %>%
      ggplot(aes(x = n_loc_rank, y = n_loc, colour = guild)) 
  } else {
    p <- freq_df %>%
      ggplot(aes(x = n_loc_rank, y = n_loc)) 
  }
  p +
    geom_step(size = 1) +
    scale_x_log10() +
    theme_bw() +
    scale_color_brewer(palette = "Paired", 
                       labels = c("plants", "pollinators")) +
    labs(x = "rank", 
         y = "# locations") +
    theme(legend.position = c(0.99, 0.99), 
          legend.justification = c(1,1))
}

sp_rank <- species_list %>%
  remove_unknown_species() %>% 
  get_sp_freq_by_group(sp_name)

sp_rank %>%
  plot_freq_step() +
  geom_hline(yintercept = 4.5, size = 0.5, linetype = 2) +
  labs(title = "number of locations per species", 
       subtitle = "species turnover in the Web of Life database", 
       caption = "dashed line indicates the (arbitrary) threshold for selecting focal species")

@@

```{r}
sp_rank %>% 
  dplyr::filter(n_loc_rank <= 10) %>%
  dplyr::select(n_loc_rank, guild, sp_name, n_loc) %>%
  dplyr::arrange(guild, n_loc_rank) %>%
  knitr::kable(col.names = c("", "guild", "species", "# locations"))
```

### Interaction turnover

@@

int_rank <- interaction_list %>%
  remove_unknown_species() %>% 
  get_int_freq_by_group(pla_name, pol_name)

int_rank %>%
  plot_freq_step() +
  geom_hline(yintercept = 2.9, size = 0.5, linetype = 2) +
  labs(title = "locations per interactions", 
       subtitle = "interactions turnover in the Web of Life database", 
       caption = "dashed line indicates the (arbitrary) threshold for selecting focal species")
```

```{r}
int_rank %>%
  dplyr::filter(n_loc_rank <= 20) %>% 
  dplyr::select(n_loc_rank, pla_name, pol_name, n_loc) %>%
  dplyr::arrange(n_loc_rank) %>%
  knitr::kable(col.names = c("", "plant", "pollinator", "# locations"), 
               caption = 'Top 20 most "widespread" interactions')
```

@@

sp_deg_rank %>%
  ggplot(aes(x = deg, y = n_loc)) +
  geom_point(aes(colour = selected, fill = selected, shape = guild)) +
  geom_label_repel(data = dplyr::filter(sp_deg_rank, selected), 
                  aes(label = sp_name), size = 3, fontface =  "italic",
                  color = RColorBrewer::brewer.pal(3, "Paired")[2]) +
  scale_x_log10() +
  # scale_y_log10() +
  scale_fill_manual(values = c(NA, RColorBrewer::brewer.pal(3, "Paired")[2])) +
  scale_color_brewer(palette = "Paired") +
  scale_shape_manual(values = c(21, 24)) +
  guides(color = guide_legend("focal species"), 
         fill = guide_legend("focal species")) +
  theme_bw() +
  theme(legend.position = c(0.02,0.98), 
        legend.justification = c(0,1), 
        legend.box = "horizontal") +
  labs(title = "species degree vs. number of locations", 
       subtitle = "there are 16 focal species, mostly generalist pollinators",
       caption = "focal species are those that have the same interaction in >= 3 locations and are present >= 5 locations", 
       y = "# locations", 
       x = "species degree")

sp_deg_rank %>%
  dplyr::filter(selected) %>% 
  dplyr::select(guild, sp_name, n_loc, deg) %>%
  dplyr::arrange(guild, deg, n_loc) %>%
  knitr::kable(col.names = c("guild", "species", "# locations", "degree"))
```

>>>>

# Interactions at the fringes, data exploration

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, dpi = 300)
```

@@

@@

We start by checking wether species and interactions occur in enough localities. 
Unknown and unindentified species are excluded from the final numbers.

If species turnover is too high the data might not be adequate to answer the research question. 
At first we set an arbitrary, but reasonable threshold of **five** locations. 
We need species to be present in at least this number of locations to have a relatively OK coverage across their range.

With this threshold of five we have 13 plants and 41 pollination species that could be candidates to be focal species.

@@

th <- 5

plot_freq_step <- function(freq_df, th = 5){
  if ("guild" %in% names(freq_df)){
    p <- freq_df %>%
      ggplot(aes(x = n_loc_rank, y = n_loc, colour = guild)) 
  } else {
    p <- freq_df %>%
      ggplot(aes(x = n_loc_rank, y = n_loc)) 
  }
  p +
    geom_step(size = 1) +
    scale_x_log10() +
    theme_bw() +
    scale_color_brewer(palette = "Paired", 
                       aesthetics = c("colour", "fill")) +
    scale_alpha_manual(values = c(0.5, 1)) +
    scale_size_manual(values = c(0.5, 1)) +
    labs(x = "rank", 
         y = "# locations") +
    theme(legend.position = c(0.98,0.98), 
          legend.justification = c(1,1), 
          legend.direction = "horizontal")
}

sp_rank <- species_list %>%
  remove_unknown_species() %>% 
  get_sp_freq_by_group(sp_name)  %>%
  dplyr::mutate(focal = n_loc >= th)

n_sp <- sp_rank %>%
  dplyr::group_by(guild, focal) %>%
  dplyr::summarise(n_sp = dplyr::n_distinct(sp_name)) %>%
  dplyr::mutate(label = paste(format(n_sp, big.mark = ","), guild)) %>%
  dplyr::group_by() %>%
  humanize()

sp_rank %>%
  dplyr::group_by() %>%
  humanize() %>%
  plot_freq_step() +
  geom_vline(data = n_sp, aes(xintercept = n_sp, 
                              linetype = focal), 
             size = 0.25, colour = "grey50") +
  geom_text(data = n_sp, 
            aes(label = label, x = n_sp, y = 15),
            size = 3.5,
            angle = 90, 
            nudge_x = -0.05, show.legend = FALSE) +
  geom_text(data = n_sp,
            aes(label = focal, x = n_sp, y = 15), 
            angle = 90, nudge_x = 0.05, colour = "grey30", 
            size = 3) +
  scale_linetype_manual(values = c(3,2)) +
  guides(size = "none", linetype = "none") +
  labs(title = "number of locations per species", 
       subtitle = "species turnover in the Web of Life database", 
       caption = "we select as focal species those in 5 or more locations")

@@

We then check the turnover of the interactions. 
We expect a large degree of turnover here, but if we are to compare interactions at the fringes of a species distributions with those at the core, we need at some of the interactions to be sampled at several locations. These  would compose the "core" of the species interactions.
At first we set an arbitrary, but reasonable threshold of **three** locations.

With this threshold we have 44 interactions that repeat over at least three locations or more. 
Specifically 22 plants and 31 pollinators are involved in these interactions.

@@

int_rank <- interaction_list %>%
  remove_unknown_species() %>% 
  get_int_freq_by_group(pla_name, pol_name) %>%
  dplyr::mutate(focal = n_loc >= 3)

n_int <- int_rank %>%
  dplyr::group_by(focal) %>%
  dplyr::summarise(n_int = dplyr::n_distinct(paste(pla_name, pol_name))) %>%
  dplyr::mutate(label = paste(format(n_int, big.mark = ","), "int."))

n_int_sp <- int_rank %>%
  dplyr::filter(focal) %>%
  dplyr::select_at(dplyr::vars(dplyr::contains("name"))) %>%
  tidyr::gather("guild", "sp_name") %>%
  dplyr::mutate(guild = substr(guild, 1,3)) %>%
  dplyr::group_by(guild) %>%
  dplyr::summarise(n_sp = dplyr::n_distinct(sp_name)) %>% {
    l <- .
    as.list(extract2(l, 2)) %>%
      `names<-`(extract2(l, 1))
  }

int_rank %>%
  plot_freq_step() +
  # geom_hline(yintercept = 2.9, size = 0.5, linetype = 2) +
  geom_vline(data = n_int, aes(xintercept = n_int, 
                                linetype = focal), 
             size = 0.25, colour = "grey50", show.legend = F) +
    scale_linetype_manual(values = c(3,2)) +
  annotate(geom = "text", x = 46, y = 3.5, label = paste(n_int_sp[[1]], "plants and", n_int_sp[[2]], "pollinators with the\nsame int. in more than 3 locations"), size = 3, hjust = 0, colour = "grey30") +
  labs(title = "locations per interactions", 
       subtitle = "interactions turnover in the Web of Life database", 
       caption = "we select as focal species those with the same interaction in 3 or more locations")
```

```{r, eval = F}
selected <- int_rank %>%
  dplyr::select(pla_name, pol_name, n_loc, focal) %>% 
  tidyr::gather("guild", "sp_name", pla_name, pol_name) %>% 
  dplyr::mutate(guild = substr(guild, 1,3),
                focal_int = TRUE) %>% 
  dplyr::group_by(guild, sp_name) %>%
  dplyr::summarise(n_loc_int = mean(n_loc), 
                   n_loc_max = max(n_loc),
                   focal_int = any(focal), 
                   n = n()) %>% 
  dplyr::distinct() %>%
  dplyr::left_join(sp_rank)

selected %>% 
  # dplyr::filter(focal, focal_int) %>%
  ggplot(aes(x = n_loc, y = n_loc_int)) +
  geom_point(aes(colour = log(n))) +
  geom_label_repel(data = dplyr::filter(selected, focal, focal_int),
                  aes(label = sp_name), size = 3, fontface =  "italic",
                  color = RColorBrewer::brewer.pal(3, "Paired")[2]) 
```

@@

median_deg <- sp_deg_rank %>%
  dplyr::group_by(selected, guild) %>%
  dplyr::summarise(deg = median(deg)) %>%
  dplyr::mutate(focal = selected) %>%
  dplyr::group_by() %>%
  humanize()

sp_deg_rank %>%
  dplyr::group_by() %>%
  humanize() %>%
  ggplot(aes(x = deg, y = n_loc)) +
  geom_vline(data = median_deg, aes(xintercept = deg, colour = selected), 
             size = 0.25, 
             linetype = 2) +
  geom_text(data = median_deg, 
            aes(y = 20, 
                label = paste("median", guild, "-", deg), 
                color = selected),
            nudge_x = -0.05,
            angle = 90, 
            size = 3) +
  geom_point(aes(colour = selected, fill = selected, shape = guild)) +
  geom_label_repel(data = dplyr::filter(sp_deg_rank, selected), 
                  aes(label = sp_name), size = 3, fontface =  "italic",
                  color = RColorBrewer::brewer.pal(3, "Paired")[2]) +
  scale_x_log10() +
  # scale_y_log10() +
  scale_fill_manual(values = c(NA, RColorBrewer::brewer.pal(3, "Paired")[2])) +
  scale_color_brewer(palette = "Paired") +
  scale_shape_manual(values = c(21, 24)) +
  guides(color = guide_legend("focal species"), 
         fill = guide_legend("focal species")) +
  theme_bw() +
  theme(legend.position = c(0.02,0.98), 
        legend.justification = c(0,1), 
        legend.box = "horizontal") +
  labs(title = "species degree vs. number of locations", 
       subtitle = "there are 16 focal species, mostly generalist pollinators",
       caption = "focal species are those that have the same interaction in >= 3 locations and are present >= 5 locations", 
       y = "# locations", 
       x = "species degree")
```
